### 2017-12-28 doc/manuscript.tex

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
    \setmainfont[]{Helvetica}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Biogeography \& environmental conditions shape bacteriophage-bacteria networks across the human microbiome},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\usepackage{setspace}
\doublespacing
\usepackage[vmargin=1in,hmargin=1in]{geometry}
\usepackage{lineno}
\linenumbers
\usepackage{parskip}
\setlength{\parskip}{7.5pt}
\exhyphenpenalty=10000
\hyphenpenalty=10000

\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

\title{Biogeography \& environmental conditions shape bacteriophage-bacteria
networks across the human microbiome}
    \usepackage{authblk}
    \font\myfont=Helvetica at 12pt
    \font\nextfont=Helvetica at 10pt
                        \author[1]{\myfont Geoffrey D Hannigan}
                    \author[2]{\myfont Melissa B Duhaime}
                    \author[3]{\myfont Danai Koutra}
                    \author[1,*]{\myfont Patrick D Schloss}
                            \affil[1]{\nextfont Department of Microbiology \& Immunology, University of Michigan, Ann
Arbor, Michigan, 48109, USA}
                    \affil[2]{\nextfont Department of Ecology \& Evolutionary Biology, University of Michigan,
Ann Arbor, Michigan, 48109, USA}
                    \affil[3]{\nextfont Department of Computer Science, University of Michigan, Ann Arbor,
Michigan, 48109, USA}
                    \affil[*]{\nextfont To whom correspondence may be addressed.}
            \date{}

\begin{document}
\maketitle

~

~

~

\textbf{\emph{Short Title}}: Network Diversity of the Healthy Human
Microbiome

\textbf{\emph{Corresponding Author Information}}\\
Patrick D Schloss, PhD\\
1150 W Medical Center Dr.~1526 MSRB I\\
Ann Arbor, Michigan 48109\\
Phone: (734) 647-5801\\
Email: pschloss@umich.edu

\newpage

\section{Abstract}\label{abstract}

Viruses and bacteria are critical components of the human microbiome and
play important roles in health and disease. Most previous work has
relied on studying bacteria and viruses independently, thereby reducing
them to two separate communities. Such approaches are unable to capture
how these microbial communities interact, such as through processes that
maintain community robustness or allow phage-host populations to
co-evolve. We implemented a network-based analytical approach to
describe phage-bacteria network diversity throughout the human body. We
built these community networks using a machine learning algorithm to
predict which phages could infect which bacteria in a given microbiome.
Our algorithm was applied to paired viral and bacterial metagenomic
sequence sets from three previously published human cohorts. We
organized the predicted interactions into networks that allowed us to
evaluate phage-bacteria connectedness across the human body. We observed
evidence that gut and skin network structures were person-specific and
not conserved among cohabitating family members. High-fat diets appeared
to be associated with less connected networks. Network structure
differed between skin sites, with those exposed to the external
environment being less connected and likely more susceptible to network
degradation by microbial extinction events. This study quantified and
contrasted the diversity of virome-microbiome networks across the human
body and illustrated how environmental factors may influence
phage-bacteria interactive dynamics. This work provides a baseline for
future studies to better understand system perturbations, such as
disease states, through ecological networks.

\newpage

\section{Author Summary}\label{author-summary}

The human microbiome, the collection of microbial communities that
colonize the human body, is a crucial component to health and disease.
Two major components to the human microbiome are the bacterial and viral
communities. These communities have primarily been studied separately
using metrics of community composition and diversity. These approaches
have failed to capture the complex dynamics of interacting bacteria and
phage communities, which frequently share genetic information and work
together to maintain ecosystem homestatsis (e.g.~kill-the-winner
dynamics). Removal of bacteria or phage can disrupt or even collapse
those ecosystems. Relationship-based network approaches allow us to
capture this interaction information. Using this network-based approach
with three independent human cohorts, we were able to present an initial
understanding of how phage-bacteria networks differ throughout the human
body, so as to provide a baseline for future studies of how and why
microbiome networks differ in disease states.

\newpage

\section{Introduction}\label{introduction}

Viruses and bacteria are critical components of the human microbiome and
play important roles in health and disease. Bacterial communities have
been associated with disease states, including a range of skin
conditions {[}1{]}, acute and chronic wound healing conditions
{[}2,3{]}, and gastrointestinal diseases, such as inflammatory bowel
disease {[}4,5{]}, \emph{Clostridium difficile} infections {[}6{]} and
colorectal cancer {[}7,8{]}. Altered human viromes (virus communities
consisting primarily of bacteriophages) also have been associated with
diseases and perturbations, including inflammatory bowel disease
{[}5,9{]}, periodontal disease {[}10{]}, spread of antibiotic resistance
{[}11{]}, and others {[}12--17{]}. Viruses act in concert with their
microbial hosts as a single ecological community {[}18{]}. Viruses
influence their living microbial host communities through processes
including lysis, host gene expression modulation {[}19{]}, influence on
evolutionary processes such as horizontal gene transfer {[}22{]} or
antagonistic co-evolution {[}26{]}, and alteration of ecosystem
processes and elemental stoichiometry {[}27{]}.

Previous human microbiome work has focused on bacterial and viral
communities, but have reduced them to two separate communities by
studying them independently {[}5,9,10,12--17{]}. This approach fails to
capture the complex dynamics of interacting bacteria and phage
communities, which frequently share genetic information and work
together to maintain ecosystem structure (e.g.~kill-the-winner dynamics
that prevent domination by a single bacterium). Removal of bacteria or
phages can disrupt or even collapse those ecosystems {[}18,28--37{]}.
Integrating these datasets as relationship-based networks allow us to
capture this complex interaction information. Studying such
bacteria-phage interactions through community-wide networks built from
inferred relationships begins to provide us with insights into the
drivers of human microbiome diversity across body sites and enable the
study of human microbiome network dynamics overall.

In this study, we characterized human-associated bacterial and phage
communities by their inferred relationships using three published paired
virus and bacteria-dominated whole community metagenomic datasets
{[}13,14,38,39{]}. We leveraged machine learning and graph theory
techniques to establish and explore the human bacteria-phage network
diversity therein. This approach built upon previous large-scale
phage-bacteria network analyses by inferring interactions from
metagenomic datasets, rather than culture-dependent data {[}33{]}, which
is limited in the scale of possible experiments and analyses. We
implemented an adapted metagenomic interaction inference model that made
some improvements upon previous phage-host interaction prediction
models. Previous approaches have utilized a variety of techniques, such
as linear models that were used to predict bacteria-phage co-occurrence
using taxonomic assignments {[}40{]}, and nucleotide similarity models
that were applied to both whole virus genomes {[}41{]} and clusters of
whole and partial virus genomes {[}42{]}. Our approach uniquely included
protein interaction data and was validated based on experimentally
determined positive and negative interactions (i.e.~who does and does
not infect whom). We built on previous modeling work as a means to our
ends, and focused on the biological insights we could gain instead of
building a superior model and presenting our work as a toolkit. We
therefore did not focus on extensive benchmarking against other existing
models {[}41,43--45{]}. Through this approach we were able to provide an
initial basic understanding of the network dynamics associated with
phage and bacterial communities on and in the human body. By building
and utilizing a microbiome network, we found that different people, body
sites, and anatomical locations not only support distinct microbiome
membership and diversity {[}13,14,38,39,46--48{]}, but also support
ecological communities with distinct communication structures and
robustness to network degradation by extinction events. Through an
improved understanding of network structures across the human body, we
aim to empower future studies to investigate how these communities
dynamics are influenced by disease states and the overall impact they
may have on human health.

\section{Results}\label{results}

\subsection{Cohort Curation and Sample
Processing}\label{cohort-curation-and-sample-processing}

We studied the differences in virus-bacteria interaction networks across
healthy human bodies by leveraging previously published shotgun sequence
datasets of purified viral metagenomes (viromes) paired with
bacteria-dominated whole community metagenomes. Our study contained
three datasets that explored the impact of diet on the healthy human gut
virome {[}14{]}, the impact of anatomical location on the healthy human
skin virome {[}13{]}, and the viromes of monozygotic twins and their
mothers {[}38,39{]}. We selected these datasets because their virome
samples were subjected to virus-like particle (VLP) purification, which
removed contaminating DNA from human cells, bacteria, etc. To this end,
the publishing authors employed combinations of filtration,
chloroform/DNase treatment, and cesium chloride gradients to eliminate
organismal DNA (e.g.~bacteria, human, fungi, etc) and thereby allow for
direct assessment of both the extracellular and fully-assembled
intracellular virome \textbf{(Supplemental Figure \ref{SequenceStats}
A-B)} {[}14,39{]}. Each research group reported quality control measures
to ensure the purity of the virome sequence datasets, using both
computational and molecular techniques (e.g.~16S rRNA gene qPCR)
\textbf{(Table S1)}. These reports confirmed that the virome libraries
consisted of highly purified virus genomic DNA.

The bacterial and viral sequences from these studies were quality
filtered and assembled into contigs. We further grouped the related
bacterial and phage contigs into operationally defined units based on
their k-mer frequencies and co-abundance patterns, similar to previous
reports \textbf{(Supplemental Figure \ref{ContigStats} -
\ref{ClusterStats})} {[}42{]}. This was done both for dimensionality
reduction and to prevent inflation of node counts by using contigs which
are expected to represent multiple fragments from the same genomes. This
was also done to create genome analogs that we could use in our
classification model which was built using genome sequences. We referred
to these operationally defined groups of related contigs as operational
genomic units (OGUs). Each OGU represented a genomically similar
sub-population of either bacteria or phages. Contig lengths within
clusters ranged between \(10^{3}\) and \(10^{5.5}\) bp
\textbf{(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})}.

While supplementing the previous virome fraction quality control
measures \textbf{(Table S1)} we found that, in light of the rigorous
purification and quality control during sample collection and
preparation, 77\% (228 / 298 operational genomic units) still had some
nucleotide similarity to a given bacterial reference genome (e-value
\textless{} \(10^{-25}\)). As most phages in these communities have been
shown to be temperate (they integrate into bacterial genomes) using
methods that included nucleotide alignments of phages to bacterial
reference genomes, we interpreted this as confirmation that the majority
of those phages were temperate and therefore shared elements with
bacterial reference genomes -- a trend previously reported {[}14{]}. We
further confirmed that while the majority of these were expected to be
temperate phages, there still remained a low level of bacterial sequence
noise, which was evident as five (1.7\%) OGUs with similar sequence
elements to the bacterial 16S rRNA gene (blastn, e-value \textless{}
\(10^{-25}\), length \textgreater{} 1,000 bp). This is in line with
previous work which has suggested that bacterial noise is an inevitable
technical issue, and we considered this noise while interpreting our
findings {[}50{]}. We also identified two OGUs as being complete, high
confidence phages using the stringent Virsorter phage identification
algorithm (class 1 confidence group) {[}51{]}.

The whole metagenomic shotgun sequence samples, which primarily
consisted of bacteria, had an average viral relative abundance of 0.4\%
\textbf{(Table S1)} {[}13,14,38,39{]}. We found that only 2\% (6 / 280
OGUs) of bacterial OGUs had significantly strong nucleotide similarity
to phage reference genomes (e-value \textless{} \(10^{-25}\))
{[}13,14,38,39{]}. No OGUs were confidently identified as lytic or
temperate phage OGUs in the bacterial dataset using the Virsorter
algorithm {[}51{]}. Together this suggests minimal bacterial OGU noise
that should be considered in the study conclusions.

\subsection{Implementing Phage-Bacteria Interaction Prediction to Build
a Community
Network}\label{implementing-phage-bacteria-interaction-prediction-to-build-a-community-network}

We predicted which phage OGUs infected which bacterial OGUs using a
random forest model trained on experimentally validated infectious
relationships from six previous publications {[}41,52--56{]}. Only
bacteria and bacteriophages were used in the model. The training set
contained 43 diverse bacterial species and 30 diverse phage strains,
including both broad and specific ranges of infection
\textbf{(Supplemental Figure \ref{ValidationOverview} A - B)}. While it
is true that there are more known phages that infect bacteria, we were
limited by the information confirming which phages do not infect certain
bacteria and attempted to keep the numbers of positive and negative
interactions similar. Phages with linear and circular genomes, as well
as ssDNA and dsDNA genomes, were included in the analysis. Because we
used DNA sequencing studies, RNA phages were not considered
\textbf{(Supplemental Figure \ref{ValidationOverview} C-D)}. This
training set included both positive relationships (a phage infects a
bacterium) and negative relationships (a phage does not infect a
bacterium). This allowed us to validate the false positive and false
negative rates associated with our candidate models, thereby building
upon previous work that only considered positive relationships {[}41{]}.
It is also worth noting that while a positive interaction is strong
evidence that the interaction exists, we must also be conscious that
negative interactions are only weak evidence for a lack of interaction
because the finding could be the result of our inability to reproduce
conditions in which those interactions occur. Altogether we decided to
maintain a balanced dataset at the cost of under-sampling the available
positive interaction information because the use of such a severely
unbalanced dataset often results in over-fit and uninformative model
training.

Four phage and bacterial genomic features were used in our random forest
model to predict infectious relationships between bacteria and phages:
1) genome nucleotide similarities, 2) gene amino acid sequence
similarities, 3) bacterial Clustered Regularly Interspaced Short
Palindromic Repeat (CRISPR) spacer sequences that target phages, and 4)
similarity of protein families associated with experimentally identified
protein-protein interactions {[}57{]}. These features were calculated
using the training set described above. We chose to utilize these
metrics that directly compare nucleotide sequences between sample phages
and bacteria, instead of relying on alignment to reference genomes or
known marker genes, because we are extrapolating our model to highly
diverse communities which we expect to diverge significantly from the
available reference genomes. The resulting random forest model was
assessed using nested cross validation, and the median area under its
receiver operating characteristic (ROC) curve was 0.788, the median
model sensitivity was 0.905, and median specificity was 0.538
\textbf{(Figure \ref{RocCurve} A)}. This balance of confident true
positives at the cost of fewer true negatives is ideal for this type of
dataset which consists of primarily positive connections
\textbf{(Supplemental Figure \ref{nestedperformance})}. Nested cross
validation of the model demonstrated that the sensitivity and
specificity of the model could vary but the overall model performance
(AUC) remained more consistent \textbf{(Supplemental Figure
\ref{crossval})}. This suggested that our model would perform with a
similar overall accuracy despite changes in sensitivity/specificity
trade-offs. The most important predictor in the model was amino acid
similarity between genes, followed by nucleotide similarity of whole
genomes \textbf{(Figure \ref{RocCurve} B)}. Protein family interactions
were moderately important to the model, and CRISPRs were largely
uninformative, due to the minimal amount of identifiable CRISPRs in the
dataset and their redundancy with the nucleotide similarity methods
\textbf{(Figure \ref{RocCurve} B)}. Approximately one third of the
training set relationships yielded no score and therefore were unable to
be assigned an interaction prediction \textbf{(Figure \ref{RocCurve}
C)}.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.85000\textwidth]{}
\caption{\textbf{Summary of Multi-Study Network Model.} \emph{(A) Median
ROC curve (dark red) used to create the microbiome-virome infection
prediction model, based on nested cross validation over 25 random
iterations. The maximum and minimum performance are shown in light red.
(B) Importance scores associated with the metrics used in the random
forest model to predict relationships between bacteria and phages. The
importance score is defined as the mean decrease in accuracy of the
model when a feature (e.g.~Pfam) is excluded. Features include the local
gene alignments between bacteria and phage genes (denoted
\texttt{blastx}; the blastx algorithm in Diamond aligner), local genome
nucleotide alignments between bacteria and phage OGUs, presence of
experimentally validated protein family domains (Pfams) between phage
and bacteria OGUs, and CRISPR targeting of bacteria toward phages
(CRISPR). (C) Proportions of samples included (gray) and excluded (red)
in the model. Samples were excluded from the model because they did not
yield any scores. Those interactions without scores were automatically
classified as not having interactions. (D) Network diameter (measure of
graph size; the greatest number of traversed vertices required between
two vertices), (E) number of vertices, and (F) number of edges
(relationships) for the total network (orange) and the individual study
sub-networks (diet study = red, skin study = yellow, twin study =
green).} \label{RocCurve}}
\end{figure}

We used our random forest model to classify the relationships between
bacteria and phage operational genomic units, which were then used to
build the interactive network. The master network, analogous to the
universal microbiome network concept previously described {[}58{]},
contained the three studies as sub-networks, which themselves each
contained sub-networks for each sample \textbf{(Supplemental Figure
\ref{NetworkDiagram})}. Metadata including study, sample ID, disease,
and OGU abundance within the community were stored in the master network
for parsing in downstream analyses \textbf{(Supplemental Figure
\ref{NetworkDiagram})}. The phage and bacteria of the master network
demonstrated both narrow broad ranges of infectious interactions
\textbf{(Supplemental Figure \ref{inthm})}. Bacterial and phage relative
abundance was recorded in each sample for each OGU and the weight of the
edge connecting those OGUs was calculated as a function of those
relative abundance values. The separate extraction of the phage and
bacterial libraries ensured a more accurate measurement of the microbial
communities, as has been outlined previously {[}59,60{]}. The master
network was highly connected and contained 72,287 infectious
relationships among 578 nodes, representing 298 phages and 280 bacteria.
Although the network was highly connected, not all relationships were
present in all samples. Relationships were weighted by the relative
abundances of their associated bacteria and phages. Like the master
network, the skin network exhibited a diameter of 4 (measure of graph
size; the greatest number of traversed vertices required between two
vertices) and included 576 (297 phages, 279 bacteria, 99.7\% total) and
72,127 (99.8\%) of the master network nodes and edges, respectively
\textbf{(Figure \ref{RocCurve} E - F)}. Additionally, the subnetworks
demonstrated narrow ranges of eccentricity across their nodes
\textbf{(Supplemental Figure \ref{eccplot})}. The phages and bacteria in
the diet and twin sample sets were more sparsely related, with the diet
study consisting of 89 (41 phages, 48 bacteria) nodes and 5,566
relationships, and the twin study containing 137 (36 phages, 101
bacteria) nodes and 17,250 relationships \textbf{(Figure \ref{RocCurve}
E - F)}. As an interesting validation measure, we identified five
(1.7\%) examples of phage OGUs which similar genomic elements to the
four previously described, broadly infectious phages isolated from Lake
Michigan (tblastx; e-value \textless{} \(10^{-25}\)) {[}61{]}.

\subsection{Role of Diet on Gut Microbiome
Connectivity}\label{role-of-diet-on-gut-microbiome-connectivity}

Diet is a major environmental factor that influences resource
availability and gut microbiome composition and diversity, including
bacteria and phages {[}14,62,63{]}. Previous work in isolated
culture-based systems has suggested that changes in nutrient
availability are associated with altered phage-bacteria network
structures {[}30{]}, although this has yet to be tested in humans. We
therefore hypothesized that a change in diet would also be associated
with a change in virome-microbiome network structure in the human gut.

We evaluated the diet-associated differences in gut virome-microbiome
network structure by quantifying how central each sample's network was
on average. We accomplished this by utilizing two common weighted
centrality metrics: degree centrality and closeness centrality. Degree
centrality, the simplest centrality metric, was defined as the number of
connections each phage made with each bacterium. We supplemented
measurements of degree centrality with measurements of closeness
centrality. Closeness centrality is a metric of how close each phage or
bacterium is to all of the other phages and bacteria in the network. A
higher closeness centrality suggests that the effects of genetic
information or altered abundance would be more impactful to all other
microbes in the system. Because these are weighted metrics, the values
are functions of both connectivity as well as community composition. A
network with higher average closeness centrality also indicates an
overall greater degree of connections, which suggests a greater
resilience against network degradation by extinction events {[}30,64{]}.
This is because more highly connected networks are less likely to
degrade into multiple smaller networks when bacteria or phages are
randomly removed {[}30,64{]}. We used this information to calculate the
average connectedness per sample, which was corrected for the maximum
potential degree of connectedness. Unfortunately our dataset was
insufficiently powered to make strong conclusions toward this
hypothesis, but this is an important observation that warrants further
investigation. Although the changes are not statistically significant
using such a small sample set, visual inspection of data suggests there
is an underlying significant difference that could be elucidated with
further sampling.

Using our small sample set, we observed that the gut microbiome network
structures associated with high-fat diets appeared less connected than
those of low-fat diets, although a greater sample size will be required
to more properly evaluate this trend \textbf{(Figure \ref{dietnetworks}
A-B)}. Five subjects were available for use, all of which had matching
bacteria and virome datasets and samples from 8-10 days following the
initiation of their diets. High-fat diets appeared to exhibit reduced
degree centrality \textbf{(Figure \ref{dietnetworks} A)}, suggesting
bacteria in high-fat environments were targeted by fewer phages and that
phage tropism was more restricted. High-fat diets also appeared to
exhibit decreased closeness centrality \textbf{(Figure
\ref{dietnetworks} B)}, indicating that bacteria and phages were more
distant from other bacteria and phages in the community. This would make
genetic transfer and altered abundance of a given phage or bacterium
less capable of impacting other bacteria and phages within the network.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.90000\textwidth]{}
\caption{\textbf{Impact of Diet and Obesity on Gut Network Structure.}
\emph{(A) Quantification of average degree centrality (number of edges
per node) and (B) closeness centrality (average distance from each node
to every other node) of gut microbiome networks of subjects limited to
exclusively high-fat or low-fat diets. Each point represents the
centrality from a human subject stool sample that was collected 8-10
days following the beginning of their defined diet. There are five
samples here, compared to the four in figure \ref{intradiv}, because one
of the was only sampled post-diet, providing us data for this analysis
but not allowing us to compare to a baseline for figure \ref{intradiv}.
(C) Quantification of average degree centrality and (D) closeness
centrality between obese and healthy adult women from the Twin gut
study. Each point represents a stool sample taken from one of the three
adult woman confirmed as obese or healthy and with matching virus and
bacteria data.} \label{dietnetworks}}
\end{figure}

In addition to diet, we observed a possible trend that obesity
influenced network structure. This was done using the three mother
samples available from the twin sample set, all of which had matching
bacteria and phage samples and confirmed BMI information. The
obesity-associated network appeared to have a higher degree centrality
\textbf{(Figure \ref{dietnetworks} C)}, but less closeness centrality
than the healthy-associated networks \textbf{(Figure \ref{dietnetworks}
D)}. These results suggested that the obesity-associated networks may be
less connected. This again comes with the caveat that this is only an
opportunistic observation using an existing sample set with too few
samples to make more substantial claims. We included this observation as
a point of interest, given the data was available.

\subsection{Individuality of Microbial
Networks}\label{individuality-of-microbial-networks}

Skin and gut community membership and diversity are highly personal,
with people remaining more similar to themselves than to other people
over time {[}13,65,66{]}. We therefore hypothesized that this personal
conservation extended to microbiome network structure. We addressed this
hypothesis by calculating the degree of dissimilarity between each
subject's network, based on phage and bacteria abundance and centrality.
We quantified phage and bacteria centrality within each sample graph
using the weighted eigenvector centrality metric. This metric defines
central phages as those that are highly abundant (\(A_{O}\) as defined
in the methods) and infect many distinct bacteria which themselves are
abundant and infected by many other phages. Similarly, bacterial
centrality was defined as those bacteria that were both abundant and
connected to numerous phages that were themselves connected to many
bacteria. We then calculated the similarity of community networks using
the weighted eigenvector centrality of all nodes between all samples.
Samples with similar network structures were interpreted as having
similar capacities for network robustness and transmitting genetic
material.

We used this network dissimilarity metric to test whether microbiome
network structures were more similar within people than between people
over time. We found that gut microbiome network structures clustered by
person (ANOSIM p-value = 0.005, R = 0.958, \textbf{Figure \ref{intradiv}
A}). Network dissimilarity within each person over the 8-10 day sampling
period was less than the average dissimilarity between that person and
others, although this difference was not statistically significant
(p-value = 0.125, \textbf{Figure \ref{intradiv} B}). Four of the five
available subjects were used because one of the subjects was not sampled
at the initial time point. The lack of statistical confidence was likely
due to the small sample size of this dataset.

Although there was evidence for gut network conservation among
individuals, we found no evidence for conservation of gut network
structures within families. The gut network structures were not more
similar within families (twins and their mothers; intrafamily) compared
to other families (other twins and mothers; inter-family) (p-value =
0.312, \textbf{Figure \ref{intradiv} C}). In addition to the gut, skin
microbiome network structure was conserved within individuals (p-value
\textless{} 0.001, \textbf{Figure \ref{intradiv} D}). This distribution
was similar when separated by anatomical sites. Most sites were
statistically significantly more conserved within individuals
\textbf{(Supplemental Figure \ref{allskin})}.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.90000\textwidth]{}
\caption{\textbf{Intrapersonal vs Interpersonal Network Dissimilarity
Across Different Human Systems.} \emph{(A) NMDS ordination illustrating
network dissimilarity between subjects over time. Each sample is colored
by subject, with each colored sample pair collected 8-10 days apart.
Dissimilarity was calculated using the Bray-Curtis metric based on
abundance weighted eigenvector centrality signatures, with a greater
distance representing greater dissimilarity in bacteria and phage
centrality and abundance. Only four subjects were included here,
compared to the five used in figure \ref{dietnetworks}, because one of
the subjects was missing the initial sampling time point and therefore
lacked temporal sampling. (B) Quantification of gut network
dissimilarity within the same subject over time (intrapersonal) and the
mean dissimilarity between the subject of interest and all other
subjects (interpersonal). The p-value is provided near the bottom of the
figure. (C) Quantification of gut network dissimilarity within subjects
from the same family (intrafamily) and the mean dissimilarity between
subjects within a family and those of other families (interfamily). Each
point represents the inter-family and intra-family dissimilarity of a
twin or mother that was sampled over time. (D) Quantification of skin
network dissimilarity within the same subject and anatomical location
over time (intrapersonal) and the mean dissimilarity between the subject
of interest and all other subjects at the same time and the same
anatomical location (interpersonal). All p-values were calculated using
a paired Wilcoxon test.}\label{intradiv}}
\end{figure}

\subsection{Network Structures Across the Human Skin
Landscape}\label{network-structures-across-the-human-skin-landscape}

Extensive work has illustrated differences in diversity and composition
of the healthy human skin microbiome between anatomical sites, including
bacteria, virus, and fungal communities {[}13,47,65{]}. These
communities vary by degree of skin moisture, oil, and environmental
exposure; features which were defined in the original publication
{[}13{]}. As viruses are known to influence microbial diversity and
community composition, we hypothesized that these differences would
still be evident after integrating the bacterial and viral datasets and
evaluating their microbe-virus network structure between anatomical
sites. To test this, we evaluated the changes in network structure
between anatomical sites within the skin dataset. The anatomical sites
and their features (e.g.~moisture \& occlusion) were defined in the
previous publication through consultation with dermatologists and
reference to previous literature {[}13{]}.

The average centrality of each sample was quantified using the weighted
eigenvector centrality metric. Intermittently moist skin sites (dynamic
sites that fluctuate between being moist and dry) were significantly
less connected than the moist and sebaceous environments (p-value
\textless{} 0.001, \textbf{Figure \ref{skinnetwork} A)}. Also, skin
sites that were occluded from the environment were much more highly
connected than those that were constantly exposed to the environment or
only intermittently occluded (p-value \textless{} 0.001, \textbf{Figure
\ref{skinnetwork} B)}.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.75000\textwidth]{}
\caption{\textbf{Impact of Skin Micro-Environment on Microbiome Network
Structure.} \emph{(A) Notched box-plot depicting differences in average
eigenvector centrality between moist, intermittently moist, and
sebaceous skin sites and (B) occluded, intermittently occluded, and
exposed sites. Notched box-plots were created using ggplot2 and show the
median (center line), the inter-quartile range (IQR; upper and lower
boxes), the highest and lowest value within 1.5 * IQR (whiskers),
outliers (dots), and the notch which provides an approximate 95\%
confidence interval as defined by 1.58 * IQR / sqrt(n). Sample sizes for
each group were: Moist = 81, Sebaceous = 56, IntMoist = 56, Occluded =
106, Exposed = 61, IntOccluded = 26. (C) NMDS ordination depicting the
differences in skin microbiome network structure between skin moisture
levels and (D) occlusion. Samples are colored by their environment and
their dissimilarity to other samples was calculated as described in
figure \ref{intradiv}. (E) The statistical differences of networks
between moisture and (F) occlusion status were quantified with an anova
and post hoc Tukey test. Cluster centroids are represented by dots and
the extended lines represent the associated 95\% confidence intervals.
Significant comparisons (p-value \textless{} 0.05) are colored in red,
and non-significant comparisons are gray.}\label{skinnetwork}}
\end{figure}

To supplement this analysis, we compared the network signatures using
the centrality dissimilarity approach described above. The dissimilarity
between samples was a function of shared relationships, degree of
centrality, and bacteria/phage abundance. When using this supplementary
approach, we found that network structures significantly clustered by
moisture, sebaceous, and intermittently moist status \textbf{(Figure
\ref{skinnetwork} C,E)}. Occluded sites were significantly different
from exposed and intermittently occluded sites, but there was no
difference between exposed and intermittently occluded sites
\textbf{(Figure \ref{skinnetwork} D,F)}. These findings provide further
support that skin microbiome network structure differs significantly
between skin sites.

\section{Discussion}\label{discussion}

Foundational work has provided a baseline understanding of the human
microbiome by characterizing bacterial and viral diversity across the
human body {[}13,14,46--48,67{]}. Here we integrated the bacterial and
viral sequence sets to offer an initial understanding of how
phage-bacteria networks differ throughout the human body, so as to
provide a baseline for future studies of how and why microbiome networks
differ in disease states. We implemented a network-based analytical
model to evaluate the basic properties of the human microbiome through
bacteria and phage relationships, instead of membership or diversity
alone. This approach enabled the application of network theory to
provide a new perspective while analyzing bacterial and viral
communities simultaneously. We utilized metrics of connectivity to model
the extent to which communities of bacteria and phages interact through
mechanisms such as horizontal gene transfer, modulated bacterial gene
expression, and alterations in abundance.

Just as gut microbiome and virome composition and diversity are
conserved in individuals {[}13,46,47,66{]}, gut and skin microbiome
network structures were conserved within individuals over time. Gut
network structure was not conserved among family members. These findings
suggested that the community properties inferred from microbiome
interaction network structures, such as robustness (meaning a more
highly connected network is more ``robust'' to network degradation
because a randomly removed bacteria or phage node is less likely to
divide or disintegrate {[}30,64{]} the overall network), the potential
for horizontal gene transfer between members, and co-evolution of
populations, were person-specific. These properties may be impacted by
personal factors ranging from the body's immune system to external
environmental conditions, such as climate and diet.

We observed evidence supporting the ability of environmental conditions
to shape gut and skin microbiome interaction network structure by
observing that diet and skin location were associated with altered
network structures. We observed evidence that diet was sufficient to
alter gut microbiome network connectivity, although this needs to be
interpreted cautiously as a case observation, due to the small sample
size. Although the available sample size was small, our findings provide
some preliminary evidence that high-fat diets are less connected than
low-fat diets and that high-fat diets may therefore lead to less robust
communities with a decreased ability for microbes to directly influence
one another. We supported this finding with the observation that obesity
may have been associated with decreased network connectivity. Together
these findings suggest the food we eat may not only impact which
microbes colonize our guts, but may also impact their interactions with
infecting phages. Further work will be required to characterize these
relationships with a larger cohort.

In addition to diet, the skin environment also influenced the microbiome
interaction network structure. Network structure differed between
environmentally exposed and occluded skin sites. The sites under greater
environmental fluctuation and exposure (the exposed and intermittently
exposed sites) were less connected and therefore were predicted to have
a higher resilience against network degradation when random nodes are
removed from the network. Likewise, intermittently moist sites
demonstrated less connectedness than the moist and sebaceous sites.
These findings agree with previous work that has shown that bacterial
community networks differ by skin environment types {[}58{]}. Together
these data suggested that body sites under greater degrees of
fluctuation harbored less connected microbiomes that are potentially
less robust to network disruption by extinction events. This points to a
link between microbiome and environmental robustness toward network
homeostasis and warrants further investigation.

While these findings take us an important step closer to understanding
the microbiome through interspecies relationships, there are caveats to
and considerations regarding our findings. First, as with most
classification models, the infection classification model developed and
applied is only as good as its training set -- in this case, the
collection of experimentally-verified positive and negative infection
data. Large-scale experimental screens for phage and bacteria infectious
interactions that report high-confidence negative interactions (i.e., no
infection) are desperately needed, as they would provide more robust
model training and improved model performance. Furthermore, just as we
have improved on previous modeling efforts, we expect that new and
creative scoring metrics will improve future performance. Other creative
and high performing models are currently being developed and the
applications of these models to community network creation will continue
to move this field forward {[}43--45{]}.

Second, although our analyses utilized the best datasets currently
available for our study, this work was done retrospectively and relied
on existing data up to seven years old. These archived datasets were
limited by the technology and costs of the time. For example, the diet
and twin studies, relied on multiple displacement amplification (MDA) in
their library preparations--an approach used to overcome the large
nucleic acids requirements typical of older sequencing library
generation protocols. It is now known that MDA results in biases in
microbial community composition {[}68{]}, as well as toward ssDNA viral
genomes {[}69,70{]}, thus rendering the resulting microbial and viral
metagenomes largely non-quantitative. Future work that employs larger
sequence datasets and that avoids the use of bias-inducing amplification
steps will build on and validate our findings, as well as inform the
design and interpretation of further studies.

Although our models demonstrated satisfactory accuracy and overall
performance, it was important to interpret our findings under the
realization that our model was not perfect. This caveat is not new to
the microbiome field, with a notable example being the use of 16S rRNA
sequencing using the V4 variable region {[}60{]}. Use of the V4 variable
region excluded detection of major bacterial members, meaning that the
findings were not able to completely describe the underlying biological
environment. Despite this caveat, skin microbiome studies provided
valuable biological insights by focusing on the community differences
between groups (e.g.~disease and healthy) which were both analyzed the
same way. Similarly, here we focused our conclusions on the differences
between the groups which were all treated the same, so that we can
minimize our dependence on a perfect predictive model.

Finally, the networks in this study were built using operational genomic
units (OGUs), which represented groups of highly similar bacteria or
phage genomes or clustered genome fragments. Similar clustering
definition and validation methods, both computational and experimental,
have been implemented in other metagenomic sequencing studies, as well
{[}42,71--73{]}. These approaches could offer yet another level of
sophistication to our network-based analyses. While this operationally
defined clustering approach allows us to study whole community networks,
our ability to make conclusions about interactions among specific phage
or bacterial species or populations is inherently limited, compared to
more focused, culture-based studies such as the work by Malki \emph{et
al} {[}61{]}. Future work must address this limitation, e.g., through
improved binning methods and deeper metagenomic shotgun sequencing, but
most importantly through an improved conceptual framing of what defines
ecologically and evolutionarily cohesive units for both phage and
bacteria {[}74{]}. Defining operational genomic units and their
taxonomic underpinnings (e.g., whether OGU clusters represent genera or
species) is an active area of work critical to the utility of this
approach. As a first step, phylogenomic analyses have been performed to
cluster cyanophage isolate genomes into informative groups using shared
gene content, average nucleotide identity of shared genes, and pairwise
differences between genomes {[}75{]}. Such population-genetic assessment
of phage evolution, coupled with the ecological implications of genome
heterogeneity, will inform how to define nodes in future iterations of
the ecological network developed here. Even though we are hesitant to
speculate on phage host ranges at low taxonomic levels in our dataset,
the data does agree with previous reports of instances of broad phage
host range {[}61,76{]}. Additionally, visualization of our dataset
interactions using the heat map approach previously used in other host
range studies, suggests a trend toward modular and nested tropism, but
we do not have the strain-level resolution that powered those previous
experimental studies.

Together our work takes an initial step towards defining bacteria-virus
interaction profiles as a characteristic of human-associated microbial
communities. This approach revealed the impacts that different human
environments (e.g., the skin and gut) can have on microbiome
connectivity. By focusing on relationships between bacterial and viral
communities, they are studied as the interacting cohorts they are,
rather than as independent entities. While our developed bacteria-phage
interaction framework is a novel conceptual advance, the microbiome also
consists of archaea and small eukaryotes, including fungi and
\emph{Demodex} mites {[}1,77{]} -- all of which can interact with human
immune cells and other non-microbial community members {[}78{]}. Future
work will build from our approach and include these additional community
members and their diverse interactions and relationships (e.g., beyond
phage-bacteria). This will result in a more robust network and a more
holistic understanding of the evolutionary and ecological processes that
drive the assembly and function of the human-associated microbiome.

\section{Materials \& Methods}\label{materials-methods}

\subsection{Code Availability}\label{code-availability}

A reproducible version of this manuscript written in R markdown and all
of the code used to obtain and process the sequencing data is available
at the following GitHub repository:

https://github.com/SchlossLab/Hannigan\_ConjunctisViribus\_ploscompbio\_2017

\subsection{Data Acquisition \& Quality
Control}\label{data-acquisition-quality-control}

Raw sequencing data and associated metadata were acquired from the NCBI
sequence read archive (SRA). Supplementary metadata were acquired from
the same SRA repositories and their associated manuscripts. The gut
virome diet study (SRA: \texttt{SRP002424}), twin virome studies (SRA:
\texttt{SRP002523}; \texttt{SRP000319}), and skin virome study (SRA:
\texttt{SRP049645}) were downloaded as \texttt{.sra} files. For clarity,
the sample sizes used for each study subset were described with the data
in the results section. Sequencing files were converted to
\texttt{fastq} format using the \texttt{fastq-dump} tool of the NCBI SRA
Toolkit (v2.2.0). Sequences were quality trimmed using the Fastx toolkit
(v0.0.14) to exclude bases with quality scores below 33 and shorter than
75 bp {[}79{]}. Paired end reads were filtered to exclude sequences
missing their corresponding pair using the
\texttt{get\_trimmed\_pairs.py} script available in the source code.

\subsection{Contig Assembly}\label{contig-assembly}

Contigs were assembled using the Megahit assembly program (v1.0.6)
{[}80{]}. A minimum contig length of 1 kb was used. Iterative k-mer
stepping began at a minimum length of 21 and progressed by 20 until 101.
All other default parameters were used.

\subsection{Contig Abundance
Calculations}\label{contig-abundance-calculations}

Contigs were concatenated into two master files prior to alignment, one
for bacterial contigs and one for phage contigs. Sample sequences were
aligned to phage or bacterial contigs using the Bowtie2 global aligner
(v2.2.1) {[}81{]}. We defined a mismatch threshold of 1 bp and seed
length of 25 bp. Sequence abundance was calculated from the Bowtie2
output using the \texttt{calculate\_abundance\_from\_sam.pl} script
available in the source code.

\subsection{Operational Genomic Unit
Binning}\label{operational-genomic-unit-binning}

Contigs often represent large fragments of genomes. In order to reduce
redundancy and the resulting artificially inflated genomic richness
within our dataset, it was important to bin contigs into operational
units based on their similarity. This approach is conceptually similar
to the clustering of related 16S rRNA sequences into operational
taxonomic units (OTUs), although here we are clustering contigs into
operational genomic units (OGUs) {[}67{]}.

Contigs were clustered using the CONCOCT algorithm (v0.4.0) {[}82{]}.
Because of our large dataset and limits in computational efficiency, we
randomly subsampled the dataset to include 25\% of all samples, and used
these to inform contig abundance within the CONCOCT algorithm. CONCOCT
was used with a maximum of 500 clusters, a k-mer length of four, a
length threshold of 1 kb, 25 iterations, and exclusion of the total
coverage variable.

OGU abundance (\(A_{O}\)) was obtained as the sum of the abundance of
each contig (\(A_{j}\)) associated with that OGU. The abundance values
were length corrected such that:

\[ { A }_{ O }=\frac { { 10 }^{ 7 }\sum _{ j=1 }^{ k }{ { A }_{ j } }  }{ \sum _{ j=1 }^{ k }{ { L }_{ j } }  } \]

Where \texttt{L} is the length of each contig \texttt{j} within the OGU.

\subsection{Operational Genomic Unit
Identification}\label{operational-genomic-unit-identification}

To confirm a lack of phage sequences in the bacterial OGU dataset, we
performed blast nucleotide alignment of the bacterial OGU representative
sequences using an e-value \textless{} \(10^{-25}\), which was stricter
than the \(10^{-10}\) threshold used in the random forest model below,
against all of the phage reference genomes available in the EMBL
database. We used a stricter threshold because we know there are genomic
similarities between bacteria and phage OGUs from the interactive model,
but we were interested in contigs with high enough similarity to
references that they may indeed be from phages. We also performed the
converse analysis of aligning phage OGU representative sequences to EMBL
bacterial reference genomes. Finally, we ran both the phage and bacteria
OGU representative sequences through the Virsorter program (1.0.3) to
identify phages (all default parameters were used), using only those in
the high confidence identification category ``class 1'' {[}51{]}.

\subsection{Open Reading Frame
Prediction}\label{open-reading-frame-prediction}

Open reading frames (ORFs) were identified using the Prodigal program
(V2.6.2) with the meta mode parameter and default settings {[}83{]}.

\subsection{Classification Model Creation and
Validation}\label{classification-model-creation-and-validation}

The classification model for predicting interactions was built using
experimentally validated bacteria-phage infections or validated lack of
infections from six studies {[}41,52--56{]}. No further reference
databases were used in our alignment procedures. Associated reference
genomes were downloaded from the European Bioinformatics Institute (see
details in source code). The model was created based on the four metrics
listed below.

The four scores were used as parameters in a random forest model to
classify bacteria and bacteriophage pairs as either having infectious
interactions or not. The classification model was built using the Caret
R package (v6.0.73) {[}84{]}. The model was trained using five-fold
cross validation with ten repeats, and the median model performance was
evaluated by training the model on 80\% of the dataset and testing
performance on the remaining 20\%. Pairs without scores were classified
as not interacting. The model was optimized using the ROC value. The
resulting model performance was plotted using the plotROC R package.

\subparagraph{Identify Bacterial CRISPRs Targeting
Phages}\label{identify-bacterial-crisprs-targeting-phages}

Clustered Regularly Interspaced Short Palindromic Repeats (CRISPRs) were
identified from bacterial genomes using the PilerCR program (v1.06)
{[}85{]}. Resulting spacer sequences were filtered to exclude spacers
shorter than 20 bp and longer than 65 bp. Spacer sequences were aligned
to the phage genomes using the nucleotide BLAST algorithm with default
parameters (v2.4.0) {[}86{]}. The mean percent identity for each
matching pair was recorded for use in our classification model.

\subparagraph{Detect Matching Prophages within Bacterial
Genomes}\label{detect-matching-prophages-within-bacterial-genomes}

Temperate bacteriophages infect and integrate into their bacterial
host's genome. We detected integrated phage elements within bacterial
genomes by aligning phage genomes to bacterial genomes using the
nucleotide BLAST algorithm and a minimum e-value of 1e-10. The resulting
bitscore of each alignment was recorded for use in our classification
model.

\subparagraph{Identify Shared Genes Between Bacteria and
Phages}\label{identify-shared-genes-between-bacteria-and-phages}

As a result of gene transfer or phage genome integration during
infection, phages may share genes with their bacterial hosts, providing
us with evidence of phage-host pairing. We identified shared genes
between bacterial and phage genomes by assessing amino acid similarity
between the genes using the Diamond protein alignment algorithm
(v0.7.11.60) {[}87{]}. The mean alignment bitscores for each genome pair
were recorded for use in our classification model.

\subparagraph{Protein - Protein
Interactions}\label{protein---protein-interactions}

The final method used for predicting infectious interactions between
bacteria and phages was the detection of pairs of genes whose proteins
are known to interact. We assigned bacterial and phage genes to protein
families by aligning them to the Pfam database using the Diamond protein
alignment algorithm. We then identified which pairs of proteins were
predicted to interact using the Pfam interaction information within the
Intact database {[}57{]}. The mean bitscores of the matches between each
pair were recorded for use in the classification model.

\subsection{Interaction Network
Construction}\label{interaction-network-construction}

The bacteria and phage operational genomic units (OGUs) were scored
using the same approach as outlined above. The infectious pairings
between bacteria and phage OGUs were classified using the random forest
model described above. The predicted infectious pairings and all
associated metadata were used to populate a graph database using Neo4j
graph database software (v2.3.1) {[}88{]}. This network was used for
downstream community analysis.

\subsection{Centrality Analysis}\label{centrality-analysis}

We quantified the centrality of graph vertices using three different
metrics, each of which provided different information graph structure.
When calculating these values, let \(G(V,E)\) be an undirected,
unweighted graph with \(|V|=n\) nodes and \(|E|=m\) edges. Also, let
\(\mathbf{A}\) be its corresponding adjacency matrix with entries
\(a_{ij} = 1\) if nodes \(V_i\) and \(V_j\) are connected via an edge,
and \(a_{ij} = 0\) otherwise.

Briefly, the \textbf{closeness centrality} of node \(V_i\) is calculated
taking the inverse of the average length of the shortest paths
(\texttt{d}) between nodes \(V_i\) and all the other nodes \(V_j\).
Mathematically, the closeness centrality of node \(V_i\) is given as:

\[ { C }_{ C }\left( { V }_{ i } \right) ={ \left( \sum _{ j=1 }^{ n }{ d\left( { V }_{ i },{ V }_{ j } \right)  }  \right)  }^{ -1 } \]

The distance between nodes (\texttt{d}) was calculated as the shortest
number of edges required to be traversed to move from one node to
another.

Intuitively, the \textbf{degree centrality} of node \(V_i\) is defined
as the number of edges that are incident to that node:

\[ { C }_{ D }\left( { V }_{ i } \right) =\sum _{ j=1 }^{ n }{ { a }_{ ij } } \]

where \(a_{ij}\) is the \(ij^{th}\) entry in the adjacency matrix
\(\mathbf{A}\).

The eigenvector centrality of node \(V_i\) is defined as the \(i^{th}\)
value in the first eigenvector of the associated adjacency matrix
\(\mathbf{A}\). Conceptually, this function results in a centrality
value that reflects the connections of the vertex, as well as the
centrality of its neighboring vertices.

The \textbf{centralization} metric was used to assess the average
centrality of each sample graph \(\mathbf{G}\). Centralization was
calculated by taking the sum of each vertex \(V_{i}\)'s centrality from
the graph maximum centrality \(C_{w}\), such that:

\[ C\left( G \right) =\frac { \sum _{ i=1 }^{ n }{ Cw -c\left( { V }_{ i } \right)  }  }{ { T } }  \]

The values were corrected for uneven graph sizes by dividing the
centralization score by the maximum theoretical centralization
(\texttt{T}) for a graph with the same number of vertices.

Degree and closeness centrality were calculated using the associated
functions within the igraph R package (v1.0.1) {[}89{]}.

\subsection{Network Relationship
Dissimilarity}\label{network-relationship-dissimilarity}

We assessed similarity between graphs by evaluating the shared
centrality of their vertices, as has been done previously. More
specifically, we calculated the dissimilarity between graphs \(G_{i}\)
and \(G_{j}\) using the Bray-Curtis dissimilarity metric and eigenvector
centrality values such that:

\[ B\left( { G }_{ i },{ G }_{ j } \right) =1-\frac { 2{ C }_{ ij } }{ { C }_{ i }+{ C }_{ j } } \]

Where \(C_{ij}\) is the sum of the lesser centrality values for those
vertices shared between graphs, and \(C_{i}\) and \(C_{j}\) are the
total number of vertices found in each graph. This allows us to
calculate the dissimilarity between graphs based on the shared
centrality values between the two graphs.

\subsection{Statistics and
Comparisons}\label{statistics-and-comparisons}

Differences in intrapersonal and interpersonal network structure
diversity, based on multivariate data, were calculated using an analysis
of similarity (ANOSIM). Statistical significance of univariate
Eigenvector centrality differences were calculated using a paired
Wilcoxon test.

Statistical significance of differences in univariate eigenvector
centrality measurements of skin virome-microbiome networks were
calculated using a pairwise Wilcoxon test, corrected for multiple
hypothesis tests using the Holm correction method. Multivariate
eigenvector centrality was measured as the mean differences between
cluster centroids, with statistical significance measured using an ANOVA
and post hoc Tukey test.

\section{Acknowledgments}\label{acknowledgments}

We thank the members of the Schloss lab for their underlying
contributions. We thank the authors of the original studies for making
their data and metadata publicly available and understandable. We also
thank the participants in the studies.

\section{Author Contributions}\label{author-contributions}

\emph{Conceptualization}: GDH, MBD, DK, PDS. \emph{Data Curation}: GDH.
\emph{Formal Analysis}: GDH. \emph{Funding Acquisition}: GDH, PDS.
\emph{Writing -- Original Draft Preparation}: GDH, PDS. \emph{Writing --
Review \& Editing}: GDH, MBD, DK, PDS.

\section{Funding Information}\label{funding-information}

GDH was supported in part by the Molecular Mechanisms in Microbial
Pathogenesis Training Program (T32 AI007528). GDH and PDS were supported
in part by funding from the NIH (P30DK034933, U19AI09087, and
U01AI124255).

\section{Competing interests}\label{competing-interests}

The authors report no conflicts of interest.

\newpage

\section{Supplemental Figure
Captions}\label{supplemental-figure-captions}

\beginsupplement

\begin{figure}[htbp]
\centering
\includegraphics[width=0.80000\textwidth]{}
\caption{\textbf{Sequencing Depth Summary.} \emph{Number of sequences
that aligned to (A) Phage and (B) Bacteria operational genomic units per
sample and colored by study.}\label{SequenceStats}}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics[width=0.50000\textwidth]{}
\caption{\textbf{Contig Summary Statistics.} \emph{Scatter plot heat map
with each hexagon representing the abundance of contigs. Contigs are
organized by length on the x-axis and the number of aligned sequences on
the y-axis.}\label{ContigStats}}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics[width=0.80000\textwidth]{}
\caption{\textbf{Operational Genomic Unit Summary Statistics.}
\emph{Scatter plot with operational genomic unit clusters organized by
average contig length within the cluster on the x-axis and the number of
contigs in the cluster on the y-axis. Operational genomic units of (A)
bacteriophages and (B) bacteria are shown.}\label{ClusterStats}}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics[width=1.00000\textwidth]{}
\caption{\textbf{Summary information of validation dataset used in the
interaction predictive model.} \emph{A) Categorical heat-map
highlighting the experimentally validated positive and negative
interactions. Only bacteria species are shown, which represent multiple
reference strains. Phages are labeled on the x-axis and bacteria are
labeled on the y-axis. B) Quantification of bacterial host strains known
to exist for each phage. C) Genome strandedness and D) linearity of the
phage reference genomes used for the
dataset.}\label{ValidationOverview}}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics[width=1.00000\textwidth]{}
\caption{\textbf{Classification Model Performance By Nested
Cross-Validation.} \emph{Box plot illustrating the median and variance
of phage-bacteria interaction prediction model. Performance was
evaluated using nested cross validation, meaning that 20\% of the
samples were randomly withheld from model training and then used to
evaluate performance. The results of 100 random iterations are shown.
Metrics include area under the curve (gray), sensitivity (red), and
specificity (tan).}\label{nestedperformance}}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics[width=1.00000\textwidth]{}
\caption{\textbf{Stable Classification Model Performance Over Random
Iterations} \emph{In addition to nested cross-validation, here we show
the results from the five-fold cross validation, in which 20\% of the
samples were randomly withheld during the training stage for model
evaluation and mtry tuning. The results of 25 random iterations are
shown. Metrics include area under the curve (red), sensitivity (green),
and specificity (blue). Dashed line highlight the random point of
0.5.}\label{crossval}}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics[width=0.50000\textwidth]{}
\caption{\textbf{Structure of the interactive network.} \emph{Metadata
relationships to samples (Phage Sample ID and Bacteria Sample ID)
included the associated time point, the study, the subject the sample
was taken from, and the associated disease. Infectious interactions were
recorded between phage and bacteria operational genomic units (OGUs).
Sequence count abundance for each OGU within each sample was also
recorded.}\label{NetworkDiagram}}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics[width=0.50000\textwidth]{}
\caption{\textbf{Heatmap of Phage-Bacteria Interaction Relationships of
Master Network.} \emph{Heatmap illustrating the ranges of infectious
interactions predicted between bacteria and bacteriophages across our
three studies. Bacterial OGUs are aligned on the vertical access, and
the bacteriophage OGUs are organized on the horizontal access. OGUs are
organized near other OGUs with similar infectious profiles, which are
further illustrated by the dendrograms. Predicted infections are tan and
predicted lacks of interactions are red.}\label{inthm}}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics[width=0.50000\textwidth]{}
\caption{\textbf{Distribution of node eccentricity across subnetworks.}
\emph{Histograms illustrating the distributions of node eccentricity
values across the subnetworks, for supplementing the node, edge, and
diameter values provided for the networks. Eccentricity of each node is
the shortest distance of that node to the furthest other node within the
graph.}\label{eccplot}}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics[width=0.75000\textwidth]{}
\caption{\textbf{Intrapersonal vs Interpersonal Dissimilarity of the
Skin.} \emph{Quantification of skin network dissimilarity within the
same subject and anatomical location over time (intrapersonal) and the
mean dissimilarity between the subject of interest and all other
subjects at the same time and the same anatomical location
(interpersonal), separated by each anatomical site (forehead {[}Fh{]},
palm {[}Pa{]}, toe web {[}Tw{]}, umbilicus {[}Um{]}, antecubital fossa
{[}Ac{]}, axilla {[}Ax{]}, and retroauricular crease {[}Ra{]}). P-value
was calculated using a paired Wilcoxon test.}\label{allskin}}
\end{figure}

\newpage

\section{Supplemental Table Captions}\label{supplemental-table-captions}

Table S1: Summary of the primary quality control measures reported in
the original publications of the viromes used in this current study.

\newpage

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-Hannigan:2013im}{}
1. Hannigan GD, Grice EA. Microbial Ecology of the Skin in the Era of
Metagenomics and Molecular Microbiology. Cold Spring Harbor Perspectives
in Medicine. 2013;3: a015362--a015362.

\hypertarget{ref-Hannigan:2014be}{}
2. Hannigan GD, Hodkinson BP, McGinnis K, Tyldsley AS, Anari JB, Horan
AD, et al. Culture-independent pilot study of microbiota colonizing open
fractures and association with severity, mechanism, location, and
complication from presentation to early outpatient follow-up. Journal of
Orthopaedic Research. 2014;32: 597--605.

\hypertarget{ref-Loesche:2016ev}{}
3. Loesche M, Gardner SE, Kalan L, Horwinski J, Zheng Q, Hodkinson BP,
et al. Temporal stability in chronic wound microbiota is associated with
poor healing. Journal of Investigative Dermatology. 2016;

\hypertarget{ref-He:2016ch}{}
4. He Q, Li X, Liu C, Su L, Xia Z, Li X, et al. Dysbiosis of the fecal
microbiota in the TNBS-induced Crohn's disease mouse model. Applied
Microbiology and Biotechnology. 2016; 1--10.

\hypertarget{ref-Norman:2015kb}{}
5. Norman JM, Handley SA, Baldridge MT, Droit L, Liu CY, Keller BC, et
al. Disease-specific alterations in the enteric virome in inflammatory
bowel disease. Cell. 2015;160: 447--460.

\hypertarget{ref-Seekatz:2016fz}{}
6. Seekatz AM, Rao K, Santhosh K, Young VB. Dynamics of the fecal
microbiome in patients with recurrent and nonrecurrent Clostridium
difficile infection. Genome medicine. 2016;8: 47.

\hypertarget{ref-Zackular:2014fba}{}
7. Zackular JP, Rogers MAM, Ruffin MT, Schloss PD. The human gut
microbiome as a screening tool for colorectal cancer. Cancer prevention
research (Philadelphia, Pa). 2014;7: 1112--1121.

\hypertarget{ref-Baxter:2014hb}{}
8. Baxter NT, Zackular JP, Chen GY, Schloss PD. Structure of the gut
microbiome following colonization with human feces determines colonic
tumor burden. Microbiome. 2014;2: 20.

\hypertarget{ref-Manrique:2016dx}{}
9. Manrique P, Bolduc B, Walk ST, Oost J van der, Vos WM de, Young MJ.
Healthy human gut phageome. Proceedings of the National Academy of
Sciences of the United States of America. 2016; 201601060.

\hypertarget{ref-Ly:2014ew}{}
10. Ly M, Abeles SR, Boehm TK, Robles-Sikisaka R, Naidu M,
Santiago-Rodriguez T, et al. Altered Oral Viral Ecology in Association
with Periodontal Disease. mBio. 2014;5: e01133--14--e01133--14.

\hypertarget{ref-Modi:2013fia}{}
11. Modi SR, Lee HH, Spina CS, Collins JJ. Antibiotic treatment expands
the resistance reservoir and ecological network of the phage metagenome.
Nature. 2013;499: 219--222.

\hypertarget{ref-Monaco:2016ita}{}
12. Monaco CL, Gootenberg DB, Zhao G, Handley SA, Ghebremichael MS, Lim
ES, et al. Altered Virome and Bacterial Microbiome in Human
Immunodeficiency Virus-Associated Acquired Immunodeficiency Syndrome.
Cell Host and Microbe. 2016;19: 311--322.

\hypertarget{ref-Hannigan:2015fz}{}
13. Hannigan GD, Meisel JS, Tyldsley AS, Zheng Q, Hodkinson BP,
SanMiguel AJ, et al. The Human Skin Double-Stranded DNA Virome:
Topographical and Temporal Diversity, Genetic Enrichment, and Dynamic
Associations with the Host Microbiome. mBio. 2015;6: e01578--15.

\hypertarget{ref-Minot:2011ez}{}
14. Minot S, Sinha R, Chen J, Li H, Keilbaugh SA, Wu GD, et al. The
human gut virome: Inter-individual variation and dynamic response to
diet. Genome Research. 2011;21: 1616--1625.

\hypertarget{ref-SantiagoRodriguez:2015gd}{}
15. Santiago-Rodriguez TM, Ly M, Bonilla N, Pride DT. The human urine
virome in association with urinary tract infections. Frontiers in
Microbiology. 2015;6: 14.

\hypertarget{ref-Abeles:2015dy}{}
16. Abeles SR, Ly M, Santiago-Rodriguez TM, Pride DT. Effects of Long
Term Antibiotic Therapy on Human Oral and Fecal Viromes. PLOS ONE.
2015;10: e0134941.

\hypertarget{ref-Abeles:2014kj}{}
17. Abeles SR, Robles-Sikisaka R, Ly M, Lum AG, Salzman J, Boehm TK, et
al. Human oral viruses are personal, persistent and gender-consistent.
2014; 1--15.

\hypertarget{ref-Haerter:2014ii}{}
18. Haerter JO, Mitarai N, Sneppen K. Phage and bacteria support mutual
diversity in a narrowing staircase of coexistence. The ISME Journal.
2014;8: 2317--2326.

\hypertarget{ref-Lindell:2005gz}{}
19. Lindell D, Jaffe JD, Johnson ZI, Church GM, Chisholm SW.
Photosynthesis genes in marine viruses yield proteins during host
infection. Nature. 2005;438: 86--89.

\hypertarget{ref-Tyler:2013fl}{}
20. Tyler JS, Beeri K, Reynolds JL, Alteri CJ, Skinner KG, Friedman JH,
et al. Prophage induction is enhanced and required for renal disease and
lethality in an EHEC mouse model. PLoS Pathogens. 2013;9: e1003236.

\hypertarget{ref-Hargreaves:2014ja}{}
21. Hargreaves KR, Kropinski AM, Clokie MR. Bacteriophage behavioral
ecology: How phages alter their bacterial host's habits. Bacteriophage.
2014;4: e29866.

\hypertarget{ref-Moon:2015fa}{}
22. Moon BY, Park JY, Hwang SY, Robinson DA, Thomas JC, Fitzgerald JR,
et al. Phage-mediated horizontal transfer of a Staphylococcus aureus
virulence-associated genomic island. Scientific Reports. 2015;5: 9784.

\hypertarget{ref-Modi:2013fi}{}
23. Modi SR, Lee HH, Spina CS, Collins JJ. Antibiotic treatment expands
the resistance reservoir and ecological network of the phage metagenome.
Nature. 2013;499: 219--222.

\hypertarget{ref-Ogg:1981th}{}
24. Ogg JE, Timme TL, Alemohammad MM. General Transduction in Vibrio
cholerae. Infection and Immunity. 1981;31: 737--741.

\hypertarget{ref-Frost:2005dn}{}
25. Frost LS, Leplae R, Summers AO, Toussaint A. Mobile genetic
elements: the agents of open source evolution. Nature Reviews
Microbiology. 2005;3: 722--732.

\hypertarget{ref-Koskella:2014ds}{}
26. Koskella B, Brockhurst MA. Bacteria-phage coevolution as a driver of
ecological and evolutionary processes in microbial communities. FEMS
Microbiology Reviews. 2014;38: 916--931.

\hypertarget{ref-Jover:2014gq}{}
27. Jover LF, Effler TC, Buchan A, Wilhelm SW, Weitz JS. The elemental
composition of virus particles: implications for marine biogeochemical
cycles. Nature Reviews Microbiology. 2014;12: 519--528.

\hypertarget{ref-Harcombe:2005fd}{}
28. Harcombe WR, Bull JJ. Impact of phages on two-species bacterial
communities. Applied and Environmental Microbiology. 2005;71:
5254--5259.

\hypertarget{ref-Middelboe:2001fl}{}
29. Middelboe M, Hagström A, Blackburn N, Sinn B, Fischer U, Borch NH,
et al. Effects of Bacteriophages on the Population Dynamics of Four
Strains of Pelagic Marine Bacteria. Microbial Ecology. 2001;42:
395--406.

\hypertarget{ref-Poisot:2011jc}{}
30. Poisot T, Lepennetier G, Martinez E, Ramsayer J, Hochberg ME.
Resource availability affects the structure of a natural
bacteriabacteriophage community. Biology letters. 2011;7: 201--204.

\hypertarget{ref-Thompson:2012ki}{}
31. Thompson RM, Brose U, Dunne JA, Hall RO, Hladyz S, Kitching RL, et
al. Food webs: reconciling the structure and function of biodiversity.
Trends in ecology \& evolution. 2012;27: 689--697.

\hypertarget{ref-Moebus:1981kp}{}
32. Moebus K, Nattkemper H. Bacteriophage sensitivity patterns among
bacteria isolated from marine waters. Helgoländer Meeresuntersuchungen.
1981;34: 375--385.

\hypertarget{ref-Flores:2013hc}{}
33. Flores CO, Valverde S, Weitz JS. Multi-scale structure and
geographic drivers of cross-infection within marine bacteria and phages.
The ISME Journal. 2013;7: 520--532.

\hypertarget{ref-Poisot:2012fh}{}
34. Poisot T, Canard E, Mouillot D, Mouquet N, Gravel D. The
dissimilarity of species interaction networks. Ecology letters. 2012;15:
1353--1361.

\hypertarget{ref-Poisot071993}{}
35. Poisot T, Stouffer D. How ecological networks evolve. bioRxiv. 2016;

\hypertarget{ref-Flores:2011bh}{}
36. Flores CO, Meyer JR, Valverde S, Farr L, Weitz JS. Statistical
structure of host-phage interactions. Proceedings of the National
Academy of Sciences of the United States of America. 2011;108: E288--97.

\hypertarget{ref-Jover:2015ev}{}
37. Jover LF, Flores CO, Cortez MH, Weitz JS. Multiple regimes of robust
patterns between network structure and biodiversity. Scientific Reports.
2015;5: 17856.

\hypertarget{ref-Reyes:2010cwa}{}
38. Reyes A, Haynes M, Hanson N, Angly FE, Heath AC, Rohwer F, et al.
Viruses in the faecal microbiota of monozygotic twins and their mothers.
Nature. 2010;466: 334--338.

\hypertarget{ref-Turnbaugh:2009ei}{}
39. Turnbaugh PJ, Hamady M, Yatsunenko T, Cantarel BL, Duncan A, Ley RE,
et al. A core gut microbiome in obese and lean twins. Nature. 2009;457:
480--484.

\hypertarget{ref-LimaMendez:2015hw}{}
40. Lima-Mendez G, Faust K, Henry N, Decelle J, Colin S, Carcillo F, et
al. Ocean plankton. Determinants of community structure in the global
plankton interactome. Science. 2015;348: 1262073--1262073.

\hypertarget{ref-Edwards:2015iz}{}
41. Edwards RA, McNair K, Faust K, Raes J, Dutilh BE. Computational
approaches to predict bacteriophage-host relationships. FEMS
Microbiology Reviews. 2015;40: 258--272.

\hypertarget{ref-Roux:2016cc}{}
42. Roux S, Brum JR, Dutilh BE, Sunagawa S, Duhaime MB, Loy A, et al.
Ecogenomics and potential biogeochemical impacts of globally abundant
ocean viruses. Nature. 2016;537: 689--693.

\hypertarget{ref-Villarroel:2016jy}{}
43. Villarroel J, Kleinheinz KA, Jurtz VI, Zschach H, Lund O, Nielsen M,
et al. HostPhinder: A Phage Host Prediction Tool. Viruses. 2016;8: 116.

\hypertarget{ref-Galiez:2017gm}{}
44. Galiez C, Siebert M, Enault F, Vincent J, Söding J. WIsH: who is the
host? Predicting prokaryotic hosts from metagenomic phage contigs.
Bioinformatics. 2017;33: 3113--3114.

\hypertarget{ref-Ahlgren:2017gh}{}
45. Ahlgren NA, Ren J, Lu YY, Fuhrman JA, Sun F. Alignment-free
\(d_2^*\) oligonucleotide frequency dissimilarity measure improves
prediction of hosts from metagenomically-derived viral sequences.
Nucleic Acids Research. 2017;45: 39--53.

\hypertarget{ref-Grice:2009eea}{}
46. Grice EA, Kong HH, Conlan S, Deming CB, Davis J, Young AC, et al.
Topographical and Temporal Diversity of the Human Skin Microbiome.
Science. 2009;324: 1190--1192.

\hypertarget{ref-Findley:2013jf}{}
47. Findley K, Oh J, Yang J, Conlan S, Deming C, Meyer JA, et al.
Topographic diversity of fungal and bacterial communities in human skin.
Nature. 2013; 1--6.

\hypertarget{ref-Costello:2009im}{}
48. Costello EK, Lauber CL, Hamady M, Fierer N, Gordon JI, Knight R.
Bacterial community variation in human body habitats across space and
time. Science. 2009;326: 1694--1697.

\hypertarget{ref-Consortium:2012iz}{}
49. Consortium THMP. A framework for human microbiome research. Nature.
2012;486: 215--221.

\hypertarget{ref-Roux:2013cr}{}
50. Roux S, Krupovic M, Debroas D, Forterre P, Enault F. Assessment of
viral community functional potential from viral metagenomes may be
hampered by contamination with cellular sequences. Open Biology. 2013;3:
130160--130160.

\hypertarget{ref-Roux:2015dh}{}
51. Roux S, Enault F, Hurwitz BL, Sullivan MB. VirSorter: mining viral
signal from microbial genomic data. PeerJ. 2015;3: e985--20.

\hypertarget{ref-Jensen:1998vh}{}
52. Jensen EC, Schrader HS, Rieland B, Thompson TL, Lee KW, Nickerson
KW, et al. Prevalence of broad-host-range lytic bacteriophages of
Sphaerotilus natans, Escherichia coli, and Pseudomonas aeruginosa.
Applied and Environmental Microbiology. 1998;64: 575--580.

\hypertarget{ref-Malki:2015tm}{}
53. Malki K, Kula A, Bruder K, Sible E. Bacteriophages isolated from
Lake Michigan demonstrate broad host-range across several bacterial
phyla. Virology. 2015;

\hypertarget{ref-Schwarzer:2012ez}{}
54. Schwarzer D, Buettner FFR, Browning C, Nazarov S, Rabsch W, Bethe A,
et al. A multivalent adsorption apparatus explains the broad host range
of phage phi92: a comprehensive genomic and structural analysis. Journal
of Virology. 2012;86: 10384--10398.

\hypertarget{ref-Kim:2012dh}{}
55. Kim S, Rahman M, Seol SY, Yoon SS, Kim J. Pseudomonas aeruginosa
bacteriophage PA1Ø requires type IV pili for infection and shows broad
bactericidal and biofilm removal activities. Applied and Environmental
Microbiology. 2012;78: 6380--6385.

\hypertarget{ref-Matsuzaki:1992gw}{}
56. Matsuzaki S, Tanaka S, Koga T, Kawata T. A Broad-Host-Range
Vibriophage, KVP40, Isolated from Sea Water. Microbiology and
Immunology. 1992;36: 93--97.

\hypertarget{ref-Orchard:2014hq}{}
57. Orchard S, Ammari M, Aranda B, Breuza L, Briganti L, Broackes-Carter
F, et al. The MIntAct project--IntAct as a common curation platform for
11 molecular interaction databases. Nucleic Acids Research. 2014;42:
D358--63.

\hypertarget{ref-Bashan:2016is}{}
58. Bashan A, Gibson TE, Friedman J, Carey VJ, Weiss ST, Hohmann EL, et
al. Universality of human microbial dynamics. Nature. 2016;534:
259--262.

\hypertarget{ref-Kleiner:2015kd}{}
59. Kleiner M, Hooper LV, Duerkop BA. Evaluation of methods to purify
virus-like particles for metagenomic sequencing of intestinal viromes.
BMC Genomics. 2015;16: 7.

\hypertarget{ref-Meisel:2016iu}{}
60. Meisel JS, Hannigan GD, Tyldsley AS, SanMiguel AJ, Hodkinson BP,
Zheng Q, et al. Skin microbiome surveys are strongly influenced by
experimental design. Journal of Investigative Dermatology. 2016;

\hypertarget{ref-Malki:2015dw}{}
61. Malki K, Kula A, Bruder K, Sible E, Hatzopoulos T, Steidel S, et al.
Bacteriophages isolated from Lake Michigan demonstrate broad host-range
across several bacterial phyla. Virology Journal. 2015;12: 164.

\hypertarget{ref-Turnbaugh:2009hf}{}
62. Turnbaugh PJ, Ridaura VK, Faith JJ, Rey FE, Knight R, Gordon JI. The
effect of diet on the human gut microbiome: a metagenomic analysis in
humanized gnotobiotic mice. Science Translational Medicine. 2009;1:
6ra14--6ra14.

\hypertarget{ref-David:2014cl}{}
63. David LA, Maurice CF, Carmody RN, Gootenberg DB, Button JE, Wolfe
BE, et al. Diet rapidly and reproducibly alters the human gut
microbiome. Nature. 2014;505: 559--563.

\hypertarget{ref-Jeong:2000ima}{}
64. Jeong H, Tombor B, Albert R, Oltvai ZN, Barabási AL. The large-scale
organization of metabolic networks. Nature. 2000;407: 651--654.

\hypertarget{ref-Grice:2009ee}{}
65. Grice EA, Kong HH, Conlan S, Deming CB, Davis J, Young AC, et al.
Topographical and Temporal Diversity of the Human Skin Microbiome.
Science. 2009;324: 1190--1192.

\hypertarget{ref-Minot:2013ih}{}
66. Minot S, Bryson A, Chehoud C, Wu GD, Lewis JD, Bushman FD. Rapid
evolution of the human gut virome. Proceedings of the National Academy
of Sciences of the United States of America. 2013;110: 12450--12455.

\hypertarget{ref-Schloss:2005hz}{}
67. Schloss PD, Handelsman J. Introducing DOTUR, a computer program for
defining operational taxonomic units and estimating species richness.
Applied and Environmental Microbiology. 2005;71: 1501--1506.

\hypertarget{ref-Yilmaz:2010jb}{}
68. Yilmaz S, Allgaier M, Hugenholtz P. Multiple displacement
amplification compromises quantitative analysis of metagenomes. Nature
Methods. 2010;7: 943--944.

\hypertarget{ref-Kim:2008to}{}
69. Kim KH, Chang HW, Nam YD, Roh SW. Amplification of uncultured
single-stranded DNA viruses from rice paddy soil. Applied and \ldots{}.
2008;

\hypertarget{ref-Kim:2011hp}{}
70. Kim K-H, Bae J-W. Amplification methods bias metagenomic libraries
of uncultured single-stranded and double-stranded DNA viruses. Applied
and Environmental Microbiology. 2011;77: 7663--7668.

\hypertarget{ref-Minot:2012ed}{}
71. Minot S, Wu GD, Lewis JD, Bushman FD. Conservation of gene cassettes
among diverse viruses of the human gut. PLOS ONE. 2012;7: e42342.

\hypertarget{ref-Deng:2014eb}{}
72. Deng L, Ignacio-Espinoza JC, Gregory AC, Poulos BT, Weitz JS,
Hugenholtz P, et al. Viral tagging reveals discrete populations in
Synechococcus viral genome sequence space. Nature. 2014;513: 242--245.

\hypertarget{ref-Brum:2015iaa}{}
73. Brum JR, Ignacio-Espinoza JC, Roux S, Doulcier G, Acinas SG, Alberti
A, et al. Ocean plankton. Patterns and ecological drivers of ocean viral
communities. Science. 2015;348: 1261498--1261498.

\hypertarget{ref-Polz:2006fi}{}
74. Polz MF, Hunt DE, Preheim SP, Weinreich DM. Patterns and mechanisms
of genetic and phenotypic differentiation in marine microbes.
Philosophical Transactions of the Royal Society B: Biological Sciences.
2006;361: 2009--2021.

\hypertarget{ref-Gregory:2016cg}{}
75. Gregory AC, Solonenko SA, Ignacio-Espinoza JC, LaButti K, Copeland
A, Sudek S, et al. Genomic differentiation among wild cyanophages
despite widespread horizontal gene transfer. BMC Genomics. 2016;17: 930.

\hypertarget{ref-PaezEspino:2016gi}{}
76. Paez-Espino D, Eloe-Fadrosh EA, Pavlopoulos GA, Thomas AD, Huntemann
M, Mikhailova N, et al. Uncovering Earth's virome. Nature. 2016;

\hypertarget{ref-Grice:2011gy}{}
77. Grice EA, Segre JA. The skin microbiome. Nature Reviews
Microbiology. 2011;9: 244--253.

\hypertarget{ref-Round:2009bz}{}
78. Round JL, Mazmanian SK. The gut microbiota shapes intestinal immune
responses during health and disease. Nature reviews Immunology. 2009;9:
313--323.

\hypertarget{ref-FASTXToolkit:wr}{}
79. Hannon GJ. FASTX-Toolkit.: GNU Affero General Public License.

\hypertarget{ref-Li:2016kd}{}
80. Li D, Luo R, Liu C-M, Leung C-M, Ting H-F, Sadakane K, et al.
MEGAHIT v1.0: A fast and scalable metagenome assembler driven by
advanced methodologies and community practices. METHODS. 2016;102:
3--11.

\hypertarget{ref-Langmead:2012jh}{}
81. Langmead B, Salzberg SL. Fast gapped-read alignment with Bowtie 2.
Nature Methods. 2012;9: 357--359.

\hypertarget{ref-Alneberg:2014fc}{}
82. Alneberg J, Bjarnason BS aacute ri, Bruijn I de, Schirmer M, Quick
J, Ijaz UZ, et al. Binning metagenomic contigs by coverage and
composition. Nature Methods. 2014; 1--7.

\hypertarget{ref-Hyatt:2012cy}{}
83. Hyatt D, LoCascio PF, Hauser LJ, Uberbacher EC. Gene and translation
initiation site prediction in metagenomic sequences. Bioinformatics.
2012;28: 2223--2230.

\hypertarget{ref-caretClassificatio:ux5fU2Litux5f1}{}
84. Kuhn M. caret: Classification and Regression Training.

\hypertarget{ref-Edgar:2007bh}{}
85. Edgar RC. PILER-CR: fast and accurate identification of CRISPR
repeats. BMC Bioinformatics. 2007;8: 18.

\hypertarget{ref-Camacho:2009fc}{}
86. Camacho C, Coulouris G, Avagyan V, Ma N, Papadopoulos J, Bealer K,
et al. BLAST+: architecture and applications. BMC Bioinformatics.
2009;10: 1.

\hypertarget{ref-Buchfink:2015ki}{}
87. Buchfink B, Xie C, Huson DH. Fast and sensitive protein alignment
using DIAMOND. Nature Methods. 2015;12: 59--60.

\hypertarget{ref-Neoj:Fuwr6PBN}{}
88. Neo4j.

\hypertarget{ref-Theigraphsoftware:vh}{}
89. Csardi G, Nepusz T. The igraph software package for complex network
research.

\end{document}